Describing vocalizations in young children: A big data approach through citizen science annotation

### 2021-01-05 202010_jslhr/paper.Rmd
revision in progress, knits to docx

figsintext        : no
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : yes

@@

Specifically, the segments extracted were automatically cut into clips of exactly 500ms, extracting neighboring silence when necessary, before being uploaded on the Zooniverse platform. We will call this level *chunk*, to convey that they are part of a bigger unit (the LENA^TM^-identified "segment"). To allow the recovery of the original segments at later stages of the analysis, the scripts created a metadata file with the mapping between 500ms chunks and their corresponding segment. Chunks were then uploaded on Zooniverse using Panoptes, their open-source, command-line based API for data handling [@bowyer2015panoptes].

@@

Following recommendations from the Zooniverse board, we collected five judgments per chunk, rather than three as used in the laboratory. Majority agreement is thus achieved when three out of five judgments agree (rather than two out of three).

@@

```{r fig-process, comment=F, message=F, hide = T, warning = F, echo = F, fig.cap = "Correspondance between LENA-identified segments (which can contain one or more vocalizations) and chunks. Notice that both segments and chunks are first annotated by multiple people, and these judgments are then combined into a single label based on simple majority. To compare laboratory and Zooniverse annotations at the level of segments, chunk-level judgments are combined using the hierarchy Canonical>Non-canonical>Crying>Laughing."}
knitr::include_graphics("fig_levels.key.pdf")

@@

In this section, we provide descriptive analyses of our dataset. According to lab annotators, `r table_lab["Canonical"]`% of segments were Canonical, `r table_lab["Non-Canonical"]`% Non-Canonical, `r table_lab["Laughing"]`% Laughing, and `r table_lab["Crying"]`% Crying, with the remaining `r table_lab["Junk"]`% being categorized as "Don't mark". Zooniverse data revealed a similar distribution:  `r table_zoon["Canonical"]`% Canonical, `r table_zoon["Non-Canonical"]`% Non-Canonical, `r table_zoon["Laughing"]`% Laughing, `r table_zoon["Crying"]`% Crying, `r table_zoon["Junk"]`% Junk.
Next, we inspected the relationship between age and child-level derived metrics, of which we had two: i) Linguistic Proportion = (Canonical+Non-Canonical)/All vocalizations (i.e., we remove Junk), and ii) Canonical Proportion = Canonical/(Canonical+Non-Canonical) (i.e., we remove Junk, Crying, and Laughing). See Figure 2 for results. **the children with Angelman syndrome scored much lower than the typically-developing infants on the canonical measure but higher on the linguistic measure.**

@@

Next, we discuss the correspondence between citizen science classifications and the laboratory gold standard, at the level of individual segments. Results were visualized with a confusion matrix showing precision and recall (Figure 3): the diagonal elements show the number of correct segment-level classifications for each class while the off-diagonal elements show non-matching classifications.

@@

Although the classification at the segment level is only moderately accurate, what we are ultimately interested in is whether citizen scientists’ classifications are able to provide a reliable snapshot of children's individual development. We illustrate this association in Figure 3. Looking at all 20 children together, we found a strong positive correlation r(`r myn["lp","all"]`) = `r myr["lp","all"]`, CI [`r mylb["lp","all"]`,`r myhb["lp","all"]`], p=`r myp["lp","all"]`] between Linguistic Proportion by child from the Zooniverse and the lab annotators' data.  When we split by participant group, correlations remain high: for Angelman syndrome r(`r myn["lp","AS"]`) = `r myr["lp","AS"]`, CI [`r mylb["lp","AS"]`,`r myhb["lp","AS"]`], p=`r myp["lp","AS"]`]; low-risk control r(`r myn["lp","LR"]`) = `r myr["lp","LR"]`, CI [`r mylb["lp","LR"]`,`r myhb["lp","LR"]`], p=`r myp["lp","LR"]`].

@@

We investigated this hypothesis by calculating the proportion of their data labeled as "Junk" for each individual child.  There was no significant difference in the proportions of their data labeled as "Junk" for children with Angelman syndrome (M = `r round(JunkXdiag$estimate["mean in group AngelmanSyndrome"],3)`) compared to the low-risk group (M = `r round(JunkXdiag$estimate["mean in group Low-RiskControl"],3)`): Welch's t(`r JunkXdiag$parameter`) = `r round(JunkXdiag$statistic,3)`, p = `r round(JunkXdiag$p.value,3)`. We therefore collapsed across groups for this exploratory analysis, and split the 20 children using a median split on the proportion of their data labeled as "Junk". Results were similar across these two post-hoc subgroups (see Figure 4). For Linguistic Proportion, the correlation across lab and Zooniverse data for the lower Junk group was r(`r myn["lp","lower_Junk"]`) = `r myr["lp","lower_Junk"]`, CI [`r mylb["lp","lower_Junk"]`,`r myhb["lp","lower_Junk"]`], p=`r myp["lp","lower_Junk"]`]; and for the higher Junk group it was r(`r myn["lp","higher_Junk"]`) = `r myr["lp","higher_Junk"]`, CI [`r mylb["lp","higher_Junk"]`,`r myhb["lp","higher_Junk"]`], p=`r myp["lp","higher_Junk"]`. For Canonical Proportion, the correlation across lab and Zooniverse data for the lower Junk group was r(`r myn["cp","lower_Junk"]`) = `r myr["cp","lower_Junk"]`, CI [`r mylb["cp","lower_Junk"]`,`r myhb["cp","lower_Junk"]`], p=`r myp["cp","lower_Junk"]`]; and for the higher Junk group it was r(`r myn["cp","higher_Junk"]`) = `r myr["cp","higher_Junk"]`, CI [`r mylb["cp","higher_Junk"]`,`r myhb["cp","higher_Junk"]`], p=`r myp["cp","higher_Junk"]`. Thus, it does not seem that a higher proportion of "Junk" judgments is an index of low quality data.

@@

Since all of these results are very similar to those obtained when first matching chunks to segments, we conclude that in the future this step may not be necessary. Instead, researchers can derive linguistic and Canonical Proportions directly from citizen scientists' chunk level judgments [which was in fact what @cychosz2019canonical did].

@@

In descriptive terms, we found that category frequency was similar across laboratory and citizen science annotations. Correlations between age and our two child-level metrics revealed different developmental patterns between our sample of older children with a diagnosis of Angelman syndrome and younger low-risk controls. Given that the two groups of children differ on both of these features (age and diagnosis), we cannot empirically tease apart these two differences. Nonetheless, inspection of Figure 2 suggests different interpretations for the two derived metrics. Overall, the pattern for Linguistic Proportion is compatible with a non-linear trajectory common to the two groups, whereby the Linguistic Proportion increases rapidly up to 20 months, and hovers around 90-100% thereafter. In this case, the lack of correlation with age in the Angelman syndrome group suggests that there is little systematic variance in this metric after 20 months of age. In contrast, the two groups seem to differ in their trajectory for Canonical Proportion, with rapid increases up to 20 months in the low-risk group, and slower decreases among children in the Angelman syndrome group.

>>>>

figsintext        : no
figurelist        : yes
tablelist         : no
footnotelist      : no
lineno            : yes

always_allow_html: true

@@

Specifically, the segments extracted were automatically cut into clips of exactly 500ms, extracting neighboring silence when necessary, before being uploaded on the Zooniverse platform. We will call this level *chunk*, to convey that they are part of a bigger unit (the LENA^TM^-identified "segment"). Note that segments and chunks are collectively referred to as clips because they are parts of the longer audio recording. To allow the recovery of the original segments at later stages of the analysis, the scripts created a metadata file with the mapping between 500ms chunks and their corresponding segment. Chunks were then uploaded on Zooniverse using Panoptes, their open-source, command-line based API for data handling [@bowyer2015panoptes].

@@

Following recommendations from the Zooniverse board, we collected five judgments per clip, rather than three as used in the laboratory.  Majority agreement is thus achieved when three out of five judgments agree (rather than two out of three).

@@

```{r fig-process, comment=F, message=F, hide = T, warning = F, echo = F, fig.cap = "Correspondence between LENA-identified segments (which can contain one or more vocalizations) and chunks. Segments and chunks are collectively referred to as clips because they are parts of the longer audio recording. Notice that clips are first annotated by multiple people, and these judgments are then combined into a single label based on simple majority. To compare laboratory and Zooniverse annotations at the level of segments, chunk-level judgments are combined using the hierarchy Canonical>Non-canonical>Crying>Laughing."}
knitr::include_graphics("fig_levels.key.pdf")

@@

In this section, we provide descriptive analyses of our dataset. According to lab annotators, `r table_lab["Canonical"]`% of segments were Canonical, `r table_lab["Non-Canonical"]`% Non-Canonical, `r table_lab["Laughing"]`% Laughing, and `r table_lab["Crying"]`% Crying, with the remaining `r table_lab["Junk"]`% being categorized as "Don't mark". Zooniverse data revealed a similar distribution:  `r table_zoon["Canonical"]`% Canonical, `r table_zoon["Non-Canonical"]`% Non-Canonical, `r table_zoon["Laughing"]`% Laughing, `r table_zoon["Crying"]`% Crying, `r table_zoon["Junk"]`% Junk.
Next, we inspected the relationship between age and child-level derived metrics, of which we had two: i) Linguistic Proportion = (Canonical+Non-Canonical)/All vocalizations (i.e., we remove Junk), and ii) Canonical Proportion = Canonical/(Canonical+Non-Canonical) (i.e., we remove Junk, Crying, and Laughing). See Figure 3 for results. **the children with Angelman syndrome scored much lower than the typically-developing infants on the canonical measure but higher on the linguistic measure.**

@@

Next, we discuss the correspondence between citizen science classifications and the laboratory gold standard, at the level of individual segments. Results were visualized with a confusion matrix showing precision and recall (Figure 4): the diagonal elements show the number of correct segment-level classifications for each class while the off-diagonal elements show non-matching classifications.

@@

Although the classification at the segment level is only moderately accurate, what we are ultimately interested in is whether citizen scientists’ classifications are able to provide a reliable snapshot of children's individual development. We illustrate this association in Figure 5. Looking at all 20 children together, we found a strong positive correlation r(`r myn["lp","all"]`) = `r myr["lp","all"]`, CI [`r mylb["lp","all"]`,`r myhb["lp","all"]`], p=`r myp["lp","all"]`] between Linguistic Proportion by child from the Zooniverse and the lab annotators' data.  When we split by participant group, correlations remain high: for Angelman syndrome r(`r myn["lp","AS"]`) = `r myr["lp","AS"]`, CI [`r mylb["lp","AS"]`,`r myhb["lp","AS"]`], p=`r myp["lp","AS"]`]; low-risk control r(`r myn["lp","LR"]`) = `r myr["lp","LR"]`, CI [`r mylb["lp","LR"]`,`r myhb["lp","LR"]`], p=`r myp["lp","LR"]`].

@@

We investigated this hypothesis by calculating the proportion of their data labeled as "Junk" for each individual child.  There was no significant difference in the proportions of their data labeled as "Junk" for children with Angelman syndrome (M = `r round(JunkXdiag$estimate["mean in group AngelmanSyndrome"],3)`) compared to the low-risk group (M = `r round(JunkXdiag$estimate["mean in group Low-RiskControl"],3)`): Welch's t(`r JunkXdiag$parameter`) = `r round(JunkXdiag$statistic,3)`, p = `r round(JunkXdiag$p.value,3)`. We therefore collapsed across groups for this exploratory analysis, and split the 20 children using a median split on the proportion of their data labeled as "Junk". Results were similar across these two post-hoc subgroups (see Figure 6). For Linguistic Proportion, the correlation across lab and Zooniverse data for the lower Junk group was r(`r myn["lp","lower_Junk"]`) = `r myr["lp","lower_Junk"]`, CI [`r mylb["lp","lower_Junk"]`,`r myhb["lp","lower_Junk"]`], p=`r myp["lp","lower_Junk"]`]; and for the higher Junk group it was r(`r myn["lp","higher_Junk"]`) = `r myr["lp","higher_Junk"]`, CI [`r mylb["lp","higher_Junk"]`,`r myhb["lp","higher_Junk"]`], p=`r myp["lp","higher_Junk"]`. For Canonical Proportion, the correlation across lab and Zooniverse data for the lower Junk group was r(`r myn["cp","lower_Junk"]`) = `r myr["cp","lower_Junk"]`, CI [`r mylb["cp","lower_Junk"]`,`r myhb["cp","lower_Junk"]`], p=`r myp["cp","lower_Junk"]`]; and for the higher Junk group it was r(`r myn["cp","higher_Junk"]`) = `r myr["cp","higher_Junk"]`, CI [`r mylb["cp","higher_Junk"]`,`r myhb["cp","higher_Junk"]`], p=`r myp["cp","higher_Junk"]`. Thus, it does not seem that a higher proportion of "Junk" judgments is an index of low quality data.

@@

Since all of these results are very similar to those obtained when collapsing chunk-level judgments to generate segment-level judgments, we conclude that in the future this step may not be necessary. Instead, researchers can derive linguistic and Canonical Proportions directly from citizen scientists' chunk-level judgments [which was in fact what @cychosz2019canonical did].

@@

In descriptive terms, we found that category frequency was similar across laboratory and citizen science annotations. Correlations between age and our two child-level metrics revealed different developmental patterns between our sample of older children with a diagnosis of Angelman syndrome and younger low-risk controls. Given that the two groups of children differ on both of these features (age and diagnosis), we cannot empirically tease apart these two differences. Nonetheless, inspection of Figure 3 suggests different interpretations for the two derived metrics. Overall, the pattern for Linguistic Proportion is compatible with a non-linear trajectory common to the two groups, whereby the Linguistic Proportion increases rapidly up to 20 months, and hovers around 90-100% thereafter. In this case, the lack of correlation with age in the Angelman syndrome group suggests that there is little systematic variance in this metric after 20 months of age. In contrast, the two groups seem to differ in their trajectory for Canonical Proportion, with rapid increases up to 20 months in the low-risk group, and slower decreases among children in the Angelman syndrome group.
